---
title: "Lab 1 Assigment"
author: "Victor S Phiri"
date: "12/8/2021"
output:
  word_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

# Load packages

```{r message=FALSE, include=FALSE}
library(tidyverse)
library(psych)
library(gridExtra)
library(readr)
library(pscl) # for pR2
library(lmtest) # for lrtest
library(Amelia)
library(dominanceanalysis) # for dominanceAnalysis()
library(broom)
library(knitr)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(modelsummary)
```

## Load the Titanic Dataset


```{r include=FALSE}
getwd()

df_Titanic=read_csv("https://raw.githubusercontent.com/kekecsz/SIMM61-Course-materials/main/Home_assignment/Titanic%20-%20training%20set.csv")
```

# Data screening 
```{r include=FALSE}
view(df_Titanic)
str(df_Titanic)
```

#Recod variable

```{r}
Clean_Titanic<- df_Titanic %>%
  mutate(Pclass = factor(Pclass, levels = c(1, 2, 3), labels = c('First Class', 'Second Class', 'Third Class')),
	Survived = factor(Survived, levels = c(0, 1), labels = c('No', 'Yes')),
	                 Sex=as_factor(Sex,levels=c(0,1),labels=c("female","male")))

	
	view(Clean_Titanic)
  str(Clean_Titanic)

```

##Select variables

```{r}

 Titanic_Selected <- Clean_Titanic %>% 
  select(Age,
         Sex,
         Pclass,
         Parch,
         Survived,
         SibSp)
         
```

## Checking the data

The data has 891 observations the Age column has missing values
representing 3% of the database

```{r}
str(Titanic_Selected) # it will help us to understand our variables and also identify NAs 
   
missmap(Titanic_Selected, main = "missing values vs oberved") # assess the presence of  missing values in the database  
 

```
#Drop NAs 

Remove NAs from the dataset We drop missing values, hence, remain with
714 observations. we use missmap to visualize our data again

```{r}
# create a new dataset without NAs called Final titanic data

Final_titanic_data<- Titanic_Selected %>% drop_na()

missmap(Final_titanic_data, main = "missing values vs oberved") # check if the all the missing values have been removed 

str(Final_titanic_data)
 summary(Final_titanic_data)
 
```

# Data visualization
We use graphs to understand our data 
 #number of survivors by sex 
proportion of survivors by gender 

```{r}
# survival rates 
prop.table(table(Final_titanic_data$Survived))

ggplot(Final_titanic_data, aes(x=Sex,fill=Survived))+ geom_bar(position = "dodge") + geom_text(stat='count',aes(label=..count..),position = position_dodge(0.9),vjust=-0.2) +
ylab("Number of Passengers")


```

survival overview 

```{r}
 Final_titanic_data %>%
ggplot() +
aes(x = Survived) +
  geom_bar(aes(y = (..count..)/sum(..count..), fill=factor(Survived)), stat= "count")+
  geom_text(aes(label = scales::percent(round((..count..)/sum(..count..),2)),
                y= ((..count..)/sum(..count..))), stat="count",
            vjust = -.25) +
  ggtitle("Survived") + labs(y = "percent")

```

Survival Rates by Sex

The graph show high death rate among men than women

```{r}
 ggplot(Final_titanic_data, aes(x=Sex,fill=Survived))+ geom_bar(position = "dodge") + geom_text(stat='count',aes(label=..count..),position = position_dodge(0.9),vjust=-0.2) +
ylab("Number of Passengers")

```
```{r}
ggplot(Final_titanic_data, aes(x=Pclass,fill=Survived))+ geom_bar(position = "dodge") + geom_text(stat='count',aes(label=..count..),position = position_dodge(0.9),vjust=-0.2) +
ylab("Number of Passengers") + xlab("Passenger Class")

```

 Survive by gender and passenger class 
```{r}
ggplot(Final_titanic_data, aes(x=Pclass,fill=Survived))+ geom_bar(position = "dodge") + geom_text(stat='count',aes(label=..count..),position = position_dodge(0.9),vjust=-0.2) +
ylab("Number of Passengers") + xlab("Passenger Class")
```

#### Titanic Survival Rates by Age by Passenger class 

Death toll was high among men in class three and also the rate of women
who died was higher than other classes

```{r}
 Final_titanic_data %>%
ggplot() +
aes(x = Age, fill = Survived) +
theme_dark()+
facet_wrap(Sex~Pclass)+
geom_histogram()+
labs(y="passanger count",
       tittle="Titanic Survival Rates by Age by Passanger class")
 
 
```

#### Class by Passenger class

the data shows that the survival rates were very low in class three as
compared to passenger classes(one and two)

```{r}
Final_titanic_data %>%
ggplot() +
aes(x = Pclass, fill = Survived) +
geom_bar()+
labs(y="passanger count",
       tittle="Titanic Survival Rates by Passanger class")

 
 
```

#### Examine survival rate by gender by passenger class 

#data show that there were more males that perished in the third class
than other classes, the number of females were relatively higher in
 
class the third class when compared to other classes

```{r}
Final_titanic_data %>%
ggplot() +
aes(x = Sex, fill = Survived) +
theme_bw()+
facet_wrap(~Pclass)+
geom_bar()+
labs(y="passanger count",
       tittle="Titanic Survival Rates by Gender by Passanger class") 
 
```

#### Survival rate with sibling and or spouse by class

```{r}
 
 Final_titanic_data %>%
ggplot() +
aes(x = SibSp, fill = Survived) +
theme_dark()+
facet_wrap(~Pclass)+
geom_bar()+
labs(y="passanger count",
       tittle="Titanic Survival Rates by Gender by Passanger class")
 
 
```

#### Titanic Survival chat

Passenger class by Parch There was a lower chance of survival for those
who were with family in the third class

```{r}
Final_titanic_data%>%
ggplot() +
aes(x = Parch, fill = Survived) +
theme_classic()+
facet_wrap(~Pclass)+
geom_bar()+
labs(y="passanger count",
       tittle="Titanic Survival Rates by Age by Passanger class by Parch")
 
```

# Design the regression model: 

The previous steps has enabled us to assess our data and come up with the
following model. Outcome valuable: Survived\
Predictors: Sex, Passenger Class(Pclass), Age, Siblings or Spouse(Sibsp)
and Parent/child(Parch)

```{r}
Tmod1 = glm(Survived~ Sex+Pclass+Age+SibSp+Parch,
family = binomial(), data = Final_titanic_data)
summary(Tmod1)
 
```

#view the coefficients
```{r}
coefficients(Tmod1)

```
# Probalility caculation

Calculate probabilities for Kate and Sue survival Kate Survival without
Leonard


```{r}

 # probility Kate without leonardo  
 
# Regression equation 
 
log_odds1 =4.36+(-2.64)*0+(-1.42)*0+(-2.65)*1+(-0.04)*20+(-0.37)*0+(-0.04)*1 # log odds 
 
odds1<- exp(log_odds1)# odds 

Kate_probability1= odds1/(1+ odds1)*100# probability 

log_odds1
odds1
Kate_probability1


```

Kate Survival with Leonard

```{r}

log_odds2=4.36+(-2.64)*0+(-1.42)*0+(-2.65)*1+(-0.04)*20+(-0.37)*1+(-0.04)*1 #log odds
Odds2=exp(log_odds2)# odds 

Kate_probability_with_Leo=  Odds2/(1+Odds2)*100 #probability 


log_odds2
Odds2
Kate_probability_with_Leo

```

Sue survival without Leonard

```{r}
#
#Odds without her father 
log_odds3 = 4.36+(-2.64)*0+(-1.42)*0+(-2.65)*1+(-0.04)*4+(-0.37)*0+(-0.04)*1
Odds3=exp(log_odds3)
Sue_Probality_without_leo=Odds3/(1+Odds3)*100

log_odds3
Odds3
Sue_Probality_without_leo 
 
```

Sue survival probability with Leonard on board

```{r}
log_odds4 = 4.36+(-2.64)*0+(-1.42)*0+(-2.65)*1+(-0.04)*4+(-0.37)*0+(-0.04)*2

Odds4=exp(log_odds4)

Probalility_Sue_Leo=Odds4/(1+ Odds4)*100 

log_odds4
Odds4
Probalility_Sue_Leo
 
 
```

# Model Performance

We use the R2 function to asses the performance of our model and the
null model. "pscl package".

The model has(-318.22 llh) which is bigger than the Null model(-482.26
llhNull). This implies that our model is better than the null model.

```{r}

  pR2(Tmod1)%>%round(2)
 

head(predict(Tmod1))

```

#### Deviance


```{r}
   # 2ll deviance 
pR2(Tmod1)["llh"] * -2 # alternative model 
 
```

#### Predicting accuracy for categorization

```{r}

Final_titanic_data = Final_titanic_data %>%
mutate(pred_Tmod1 = predict(Tmod1)) %>% 
  mutate(pred_Tmod1 = case_when(pred_Tmod1 <= 0 ~"No",
                                pred_Tmod1 > 0 ~ "Yes"))
```

#### Cording correct cases

A new variable will be added to the our final titanic dataset as shown
below;

we shall use mutate and case_when function

```{r}

 Final_titanic_data = Final_titanic_data %>%
  mutate(correct_prediction=case_when(pred_Tmod1 == Survived  ~ "correct",
                                      pred_Tmod1 != Survived ~ "incorrect"))

```

#### Check the model prediction performance

The model has 81% correct prediction rate and 19%
incorrect predictions

```{r}
 
Final_titanic_data %>%
group_by(correct_prediction) %>%
summarise(count = n()) %>%
mutate(freq = count / sum(count)) 


```

# compute the null model; 

this will be used to compare the data with the data predicted by our
model.

  

```{r}
mod_null = glm(Survived ~ 1, family = binomial(), data = Final_titanic_data)

summary(mod_null)

```


#The null model
the model show that 59% of the people died while 41% died 

```{r}

Final_titanic_data %>%
group_by(Survived) %>%
summarise(count = n()) %>%
mutate(freq = count / sum(count)) 
 
```

compare the predictions of the null model with the alternative model

```{r}

Final_titanic_data%>%
group_by(Survived, pred_Tmod1) %>%
summarize(n = n()) %>%
spread(Survived, n)


```

### check for the prediction of Survival rate by response category(Survived= Yes/No)

73% of survivors were correctly predicted. However, 27% were wrongly
predicted.  

```{r}

Final_titanic_data %>%
filter(Survived == "Yes") %>%
group_by(correct_prediction) %>%
  summarise(count = n()) %>%
  mutate(freq = count / sum(count)) %>%  
mutate(freq = count / sum(count))


```

#### Predictions for those who died

our model had 86% correct prediction and 14% incorrect predictions

```{r}


Final_titanic_data %>%
filter(Survived == "No") %>%
group_by(correct_prediction) %>%
summarise(count = n()) %>% mutate(freq = count / sum(count)) 

```

#model fit 

##The AIC for the model is 650.46 which is lower than 966.52 for the null model as shown in the output below 

```{r}

 AIC(mod_null, Tmod1)
 
```

#Dominance Analysis 

```{r message=FALSE, results='asis'}

Dominance_TMod1<-dominanceAnalysis(Tmod1)

contributionByLevel(Dominance_TMod1, fit.functions="r2.m")

```
# Average contribution of predictors 

We conclude from our findings that Sex(0.21) is the main predictor of Survival outcome, followed by Passenger class(0.11), Age(0.02), Sibling or spouse(0.01), and parent or child(0.00)

note(above figures have been round to two decimal places)

```{r}

averageContribution(Dominance_TMod1,fit.functions = "r2.m")
   

```

Graph conditional dominance Analysis 


```{r}

plot(Dominance_TMod1, which.graph ="conditional",fit.function = "r2.m")



```


# Graph general Dominace 
```{r}
plot(Dominance_TMod1, which.graph ="general",fit.function = "r2.m")


```
Presentation of summary table
We look for summary tables for the report 

```{r}
tab_model(Tmod1,CSS = list(css.depvarhead = '+color: red;',css.firsttablecol = 'font-weight: bold;',css.summary = 'color: blue;',css.est=TRUE))
```






